---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can edit [this `.rmd` file](https://github.com/UBC-STAT/STAT545.github.io/blob/main/content/mini-data-analysis/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome to the rest of your mini data analysis project!

In Milestone 1, you explored your data. and came up with research questions. This time, we will finish up our mini data analysis and obtain results for your data by: 

- Making summary tables and graphs 
- Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

We will also explore more in depth the concept of *tidy data.* 

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class, and indicate *why* you had to deviate. Feel free to contact the instructor in these cases. 

# Instructions

**To complete this milestone**, edit [this very `.Rmd` file](https://github.com/UBC-STAT/STAT545.github.io/blob/main/content/mini-data-analysis/mini-project-2.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work here--->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to your mini-analysis GitHub repository, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 50 points: 45 for your analysis, and 5 for overall reproducibility, cleanliness, and coherence of the Github submission. 

**Research Questions**: In Milestone 1, you chose four research questions to focus on. Wherever realistic, your work in this milestone should relate to these research questions whenever we ask for justification behind your work. In the case that some tasks in this milestone don't align well with one of your research questions, feel free to discuss your results in the context of a different research question.

# Learning Objectives

By the end of this milestone, you should:

-   Understand what *tidy* data is, and how to create it using `tidyr`.
-   Generate a reproducible and clear report using R Markdown.
-   Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

# Setup

Begin by loading your data and the tidyverse package below:

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
library(tidyverse)
```

# Task 1: Process and summarize your data 

From Milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions. 

### 1.1 (1 point) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1. what's the relationship between number of unit and number of apartment built in a year. Because the built year distribution looks similar to number of unit per year.

2. What kind of apartment intend to missing more value than others?

3. there is one unit has 4000+ unit what make it so unic? Is there any special feature reflected on other paramete?

4. Whats the realtionship between AC coverage rate and their built year. Will that be similar to number of unit distribution?
<!----------------------------------------------------------------------------->

Here, we will investigate your data using various data manipulation and graphing functions.

### 1.2 (8 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

Also make sure that you're using dplyr and ggplot2 rather than base R. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2.  Compute the number of observations for at least one of your *categorical variables*. Do not use the function `table()`!
3.  Create a categorical variable with 3 or more groups from an existing *numerical variable*. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Compute the proportion and counts in each category of one *categorical variable* across the groups of another *categorical variable* from your data. Do not use the function `table()`!

**Graphing:**

6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
7. Make a graph where it makes sense to customize the alpha transparency.

Using variables and/or tables you made in one of the "Summarizing" tasks: 

8. Create a graph that has at least two geom layers. 
9. Create 3 histograms, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
Research Question1: 
 
what's the relationship between number of unit and number of apartment built in a year. Because the built year distribution looks similar to number of unit per year.

Task 2:

Compute the number of observations for at least one of your *categorical variables*. Do not use the function `table()`!

Note: So here I used the count of year_built. The reason why I take year_built here as a "categorical variables" is because the year make no sense if you add them up or doing any math on it so even its saved as a numver, it more represents a categorical variables just to distinguish like 1901 and 1900. And I count the year_built because this form the key variable in my research quetion which is "number of apartment built in a year". So all the following research is based on this.

```{r}
year_summary <- apt_buildings %>%
  filter(!is.na(year_built)) %>%
  group_by(year_built) %>%
  summarise(
    total_units = sum(no_of_units, na.rm = TRUE),
    n_buildings = n(),
    .groups = "drop"
  )
print(year_summary)

```
Research Question1: 
 
what's the relationship between number of unit and number of apartment built in a year. Because the built year distribution looks similar to number of unit per year.

Task 9:

Create 3 histograms, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Note: So I draw 3 bin size for 2 variables I'm interested in(Totally 6 graph). This is actually following the "Mask" thought I mentioned in task3 above. Here we have a larger bin size is works as those "Mask" that I wanna use for ignoring tiny difference. From the result shown I would say the best mask length is 5. Cause it do minimal those tiny difference but also is not fixing some meaningful information like the 10 graph.
```{r}
year_summary <- apt_buildings %>%
  filter(!is.na(year_built)) %>%
  group_by(year_built) %>%
  summarise(
    total_units = sum(no_of_units, na.rm = TRUE),
    n_buildings = n(),
    .groups = "drop"
  )

bin_widths <- c(2, 5, 10)

for (bw in bin_widths) {
  binned <- year_summary %>%
    mutate(bin = floor(year_built / bw) * bw) %>%
    group_by(bin) %>%
    summarise(
      total_units = sum(total_units, na.rm = TRUE),
      n_buildings = sum(n_buildings, na.rm = TRUE),
      .groups = "drop"
    )
  
  p1 <- ggplot(binned, aes(x = bin, y = total_units)) +
    geom_col() +
    labs(title = paste("Total Units Built (binwidth =", bw, ")"),
         x = "Year Bin Start", y = "Total Units")
  print(p1)
  
  p2 <- ggplot(binned, aes(x = bin, y = n_buildings)) +
    geom_col() +
    labs(title = paste("Buildings Built (binwidth =", bw, ")"),
         x = "Year Bin Start", y = "Building Count")
  print(p2)
}
```

Research Question 2: 

What kind of apartment intend to missing more value than others?

Task4:

Compute the proportion and counts in each category of one *categorical variable* across the groups of another *categorical variable* from your data. Do not use the function `table()`!

Note: So Here its actually the combination of task4 and 2 new created variable called missing _category and unit_group. So here we set a row is complete if all information is provided and set it to missing if any column is missing value. Unit group is based on no_of_unit and is has small or medium or large option. And we compute the missing or complete variables portion to each unit_group. The result shows large apartment has 26% to have complete information while medium has 4% and small has 1% chance. Which told us larger apartment tend to have higher chance to have complete information.

```{r}
apt_buildings <- apt_buildings %>%
  mutate(
    unit_group = case_when(
      no_of_units <= 10 ~ "Small",
      no_of_units > 10 & no_of_units <= 50 ~ "Medium",
      no_of_units > 50 ~ "Large",
      TRUE ~ "Unknown"
    ),
    missing_category = if_else(rowSums(is.na(.)) > 0, "Missing", "Complete")
  )

apt_buildings %>%
  group_by(unit_group, missing_category) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(unit_group) %>%
  mutate(prop = count / sum(count))
```

Research Question 2: 

What kind of apartment intend to missing more value than others?

Task7

Make a graph where it makes sense to customize the alpha transparency.

Note: 
So based on the information we get above. Here we made a transparency graph based on the huge chance of overlapping between complete dot and missing dot. Here we can find all the complete dot assembles in the right half of the picture, which shows another trend that the newer the apartment are the higher chance the complete information it has.
```{r}
ggplot(apt_buildings, aes(x = year_built, y = no_of_units, color = missing_category)) +
  geom_point(alpha = 0.4, size = 3) +
  scale_y_log10(labels = scales::comma) +
  labs(
    title = "Alpha Transparency Example: Year Built vs Number of Units",
    x = "Year Built",
    y = "Number of Units (log scale)",
    color = "Data Status"
  ) +
  theme_minimal(base_size = 13)
```

Research Question 3: 

there is one unit has 4000+ unit what make it so unic? Is there any special feature reflected on other paramete?

Task1:

Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.

Note: I tried a few times first find the min no_of_unit has 0 so I filtered out those.THen I have range, mean, median, min, and max here according to their property_type. We can see the private has largest range but common mean and has biggest 4111 unit apartment. This tells that TCHC has averagely most unit per apartment because its 146 mean units but private has biggest range while social housing has no obvious feature. this helps me understand the 4111 unic feature of its property type.
```{r}
apt_buildings %>%
  filter(!is.na(no_of_units), no_of_units > 0) %>%   
  group_by(property_type) %>%
  summarise(
    range_units = max(no_of_units, na.rm = TRUE) - min(no_of_units, na.rm = TRUE),
    mean_units = mean(no_of_units, na.rm = TRUE),
    median_units = median(no_of_units, na.rm = TRUE),
    min_units = min(no_of_units, na.rm = TRUE),
    max_units = max(no_of_units, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_units))
```

Research Question 3: 

there is one unit has 4000+ unit what make it so unic? Is there any special feature reflected on other parameter?

Task6: 

Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.

Note:
So this graph the y-axis is log axis, which help better shows the 4111 no_of_unit number in this graph or it will leaves too far from other dots in linear y axis that will make dots beblow unreadable. This also helps me see the 4111 located in ward 13 as its feature.
```{r}
ggplot(apt_buildings, aes(x = ward, y = no_of_units)) +
  geom_jitter(alpha = 0.4, color = "blue", width = 0.2) +
  scale_y_log10(labels = scales::comma) +
  labs(
    title = "Number of Units per Ward (Log Scale)",
    x = "Ward",
    y = "Number of Units (log scale)"
  ) +
  theme_minimal(base_size = 13)
```

Research Question 4: 

Whats the realtionship between AC coverage rate and their built year. Will that be similar to number of unit distribution?

Task4: 
Compute the proportion and counts in each category of one *categorical variable* across the groups of another *categorical variable* from your data. Do not use the function `table()`!

Note:
I can only connect this task with my research question when I compute another categoraical variable based on year_built here as built_group. so this gives me a view that coverage is 9%, 11%, 33% and 74% as time goes. so the newer apartment has higher coverage. It does not like the no_of_unit distribution with built_year.
```{r}
ac_summary <- apt_buildings %>%
  filter(!is.na(air_conditioning), !is.na(year_built)) %>%
  mutate(
    ac_category = if_else(air_conditioning == "NONE", "No AC", "Has AC"),
    built_group = case_when(
      year_built < 1960 ~ "Before 1960",
      year_built >= 1960 & year_built < 1980 ~ "1960–1979",
      year_built >= 1980 & year_built < 2000 ~ "1980–1999",
      year_built >= 2000 ~ "2000 and later"
    )
  ) %>%
  group_by(built_group, ac_category) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(built_group) %>%
  mutate(proportion = count / sum(count))
print(ac_summary)
```

Research Question 4: 

Whats the realtionship between AC coverage rate and their built year. Will that be similar to number of unit distribution?

Task8: 
Create a graph that has at least two geom layers.

Note:
So this is a more intuitive graph showing the porpotion goes along with time. The lines are the prportion has AC with proportion with no AC, which intuitively shows the trends answering question AC coverage goes higher as time goes its not similar to number of unit distribution. 
```{r}
ac_summary$built_group <- factor(
  ac_summary$built_group,
  levels = c("Before 1960", "1960–1979", "1980–1999", "2000 and later")
)

ac_has <- ac_summary %>% filter(ac_category == "Has AC")
ac_no  <- ac_summary %>% filter(ac_category == "No AC")

ggplot() +
  geom_line(
    data = ac_no,
    aes(x = built_group, y = proportion, group = 1, color = "No AC"),
    size = 1.2
  ) +
  geom_line(
    data = ac_has,
    aes(x = built_group, y = proportion, group = 1, color = "Has AC"),
    size = 1.2
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Proportion of Buildings With and Without AC by Year Built Group",
    x = "Year Built Group",
    y = "Proportion of Buildings",
    color = "AC Category"
  ) +
  theme_minimal(base_size = 13)
```


<!----------------------------------------------------------------------------->

### 1.3 (2 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->
So the question 1 about the similarity still keeps unclear on how similar they are because no correlation analysis is done on that. For question 2 we found the smaller apartment seems to missing more value but the unclearness is it could be analysis from another aspects like property-type. The question3 has almost no progress because none of the 2 features is really understandable for its unic level. But the question4 is almost done.

To refine my question, I think question 3 can be refined because there is no standard of "unic" and no standard to see how other feature contributes to that "unic". So a better question should be: In all the numeric feature, which of the parameter of 4111 apartment is not located in the range of standard deviation of that features.

The interesting result is on the similarity between apartment built per year and number of unit trend. Besides the similarity, they also shows a peak at 1970 but not like AC goes higher as time goes. So this may reflects there is some certain reason cause people start building more apartment and bigger apartment these days.

<!----------------------------------------------------------------------------->

# Task 2: Tidy your data 

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

-   Each row is an **observation**
-   Each column is a **variable**
-   Each cell is a **value**

### 2.1 (2 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have \>8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->
So here we print out our data below:
```{r}
print(apt_buildings)
```
We can see a row is a observation of one apartment, one column is one variable of that apartment and one cell is detailed value of that feature so it is a tidy data.
<!----------------------------------------------------------------------------->

### 2.2 (4 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->
I'm here only choose 8 column so its more easy to show the result.
```{r}
library(tidyr)

apt8 <- apt_buildings %>%
  select(id, year_built, no_of_units, air_conditioning,
         balconies, sprinkler_system, emergency_power, ward)
head(apt8)
apt_untidy <- apt8 %>%
  unite(
    col = "amenity_bundle",
    air_conditioning, balconies, sprinkler_system, emergency_power,
    sep = " | ",
    remove = TRUE
  )

head(apt_untidy)

apt_tidy_back <- apt_untidy %>%
  separate(
    col = amenity_bundle,
    into = c("air_conditioning", "balconies", "sprinkler_system", "emergency_power"),
    sep = " \\| ",
    remove = TRUE
  )

head(apt_tidy_back)

```
we can see the middle result amenity_bundle has 3 value in one cell so thats an untidy data. And it turned back when we seperate them by |. 
<!----------------------------------------------------------------------------->

### 2.3 (4 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in the remaining tasks:

<!-------------------------- Start your work below ---------------------------->

1. what's the relationship between number of unit and number of apartment built in a year. Because the built year distribution looks similar to number of unit per year.

2. What kind of apartment intend to missing more value than others?

<!----------------------------------------------------------------------------->

Explain your decision for choosing the above two research questions.

<!--------------------------- Start your work below --------------------------->
Because the question1 and 2 is still unclear comparing to question 4 which is already solved and still solvable comparing to question 3 which is ambiguous on its standard.
<!----------------------------------------------------------------------------->

Now, try to choose a version of your data that you think will be appropriate to answer these 2 questions. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelevant columns, etc.).

(If it makes more sense, then you can make/pick two versions of your data, one for each research question.) 

<!--------------------------- Start your work below --------------------------->

# Task 3: Modelling

## 3.0 (no points)

Pick a research question from 1.2, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: what's the relationship between number of unit and number of apartment built in a year ?

**Variable of interest**: number of unit (and number of apartment built per year)

<!----------------------------------------------------------------------------->

## 3.1 (3 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545.

-   **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit.

    -   You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    -   You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    -   You could use `lm()` to test for significance of regression coefficients.

<!-------------------------- Start your work below ---------------------------->
```{r}
year_summary <- apt_buildings %>%
  filter(!is.na(year_built), !is.na(no_of_units)) %>%
  group_by(year_built) %>%
  summarise(
    total_units = sum(no_of_units),
    n_buildings = n()
  )

model_units_buildings <- lm(total_units ~ n_buildings, data = year_summary)

summary(model_units_buildings)
```
I don't remember what those number means so I asked *chatgpt* and they said:
*The relationship is strong and positive — years with more buildings also have many more total units built.*

*The slope is highly significant (p < 0.001).*

*R² = 0.65 → the model explains about 65% of the variation in total units per year using just the number of buildings built.*
<!----------------------------------------------------------------------------->

## 3.2 (3 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value.

-   Be sure to indicate in writing what you chose to produce.
-   Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
-   Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible.

<!-------------------------- Start your work below ---------------------------->
```{r}
library(broom)


year_summary <- apt_buildings %>%
  filter(!is.na(year_built), !is.na(no_of_units)) %>%
  group_by(year_built) %>%
  summarise(
    total_units = sum(no_of_units),
    n_buildings = n()
  )

model_units_buildings <- lm(total_units ~ n_buildings, data = year_summary)

model_tidy <- tidy(model_units_buildings)

result <- model_tidy %>%
  filter(term == "n_buildings") %>%
  select(term, estimate, p.value)

print(result)

```
This means each building built averagely will lead to 78 unit built and its highly relavent.
<!----------------------------------------------------------------------------->

# Task 4: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there.

## 4.1 (3 points)

Take a summary table that you made from Task 1, and write it as a csv file in your `output` folder. Use the `here::here()` function.

-   **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
-   **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file.

<!-------------------------- Start your work below ---------------------------->
```{r}
library(readr)


write_csv(
  year_summary,
  here::here("output", "year_summary.csv")
)
```
<!----------------------------------------------------------------------------->

## 4.2 (3 points)

Write your model object from Task 3 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`.

-   The same robustness and reproducibility criteria as in 4.1 apply here.

<!-------------------------- Start your work below ---------------------------->
The code below saved and read the untidy model in task3 and compared if its equal.
```{r}
library(here)
dir.create(here("output"), showWarnings = FALSE)
saveRDS(model_units_buildings, file = here("output", "model_units_buildings.rds"))
model_units_buildings_loaded <- readRDS(here("output", "model_units_buildings.rds"))
identical(class(model_units_buildings_loaded), class(model_units_buildings))
all.equal(coef(model_units_buildings_loaded), coef(model_units_buildings))
```
<!----------------------------------------------------------------------------->

# Overall Reproducibility/Cleanliness/Coherence Checklist 

Here are the criteria we're looking for.

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. 

The README file should still satisfy the criteria from the last milestone, i.e. it has been updated to match the changes to the repository made in this milestone. 

## File and folder structure (1 points)

You should have at least three folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else.

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (1 point)

All output is recent and relevant:

-   All Rmd files have been `knit`ted to their output md files. 
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant.

## Tagged release (0.5 point)

You've tagged a release for Milestone 2. 

